Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-02-22 ms.tex
double-checked Bondi
Fixing Ia scaling

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other properties of the
galactic nucleus informs key questions related to the co-evolution of SMBHs and their host galaxies with cosmic time (e.g.~\citealt{Kormendy&Ho13}; \citealt{Heckman&Best14}).  SMBH growth in the low redshift universe is dominated by low mass black holes, $M_{\bullet} \lesssim 10^{8}M_{\odot}$ (e.g.~\citealt{Heckman+04}), a fact which is often attributed to the general trend of `cosmic down-sizing' resulting from hierarchical structure growth (e.g, \citealt{Gallo+08}).  However, the physical processes involved in the growth of loss mass blacks may be distinct from those operating at higher SMBH masses and accretion rates in quasars.  A key question here is whether SMBHs grow primarily by accreting gas fed in directly from large scales, or whether significant growth results also from internal, secular processes such as local stellar mass loss.

@@

One observes a similarity between equation (\ref{eq:mdot_gas}) and the Bondi formula (eq.~[\ref{eq:bondi}]) if $r_{\rm B}$ is replaced by $\rs$.  Indeed, when the stagnation radius resides interior to the sphere of influence we find that (eq.~[\ref{eq:stag_simple}])
\begin{align}
  \rs\approx\frac{7}{2}\frac{G \Mbh}{n \vwO^2} \approx \frac{7}{6n}r_{\rm B},
  \label{eq:rs_simple}
\end{align}
where the second equality makes use of equation (\ref{eq:Tanalytic}).  {\bf BDM: please check, as I found something different than what was here before}.

@@

To within a constant of order unity, equation (\ref{eq:cooling2}) also equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3n kT/2)/\dot{q}_{\rm rad}$ to the free-fall time $t_{\rm ff}$ at the stagnation radius, as can be derived using the equality 
\begin{align}
\rho|_{r_{\rm s}}=\frac{3 q\tff|_{r_{\rm s}}}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations (\ref{eq:Tanalytic}), \eqref{eq:mdot_analytic}, and~\eqref{eq:mdot_gas}.  \citet{McCourt+12} argue cooling instability develops in a hydrostatic atmosphere if $t_{\rm cool} \ll t_{\rm ff}$, so equation (\ref{eq:cooling2}) represents a good proxy for instability in this case as well.

@@

In Appendix \ref{app:windheat} we calculate the wind heating rate from stellar winds, $v_{\rm w}^{\star}$, and the mass loss parameter, $\eta$ (eq.~[\ref{eq:q}]), as a function of age, $\tau_{\star}$, of a stellar population which is formed impulsively (Fig.~\ref{fig:vwimp}).  At the earliest times ($\tau_{\star} \lesssim 10^{7}$ yr), the wind heating rate exceeds 1000 km s$^{-1}$, while much later ($\tau_{\star} \sim t_{\rm h}$) stellar wind heating dominated by main sequence winds is much lower, $v^{\star}_{\rm w} \sim 50-100 $ km s$^{-1}$.  We show in $\S\ref{sec:combined}$ that for the case of quasi-continueous star formation representative of the average star formation history of loss mass galaxies, the stellar wind heating rate produced by stellar winds can also be significant, $v_{\rm w}^{\star} \gtrsim 1000$ km s$^{-1}$.  Stellar winds thus contributes a potentially important source of both energy and mass for the CNM.

@@

Assuming that $v_{w}^{\rm Ia} \gg \sigma$ then by substituting $v_{w}^{\rm Ia}$ (eq.~\ref{eq:vIa}) into equation (\ref{eq:rs_simple}) for the stagnation radius, we find that
\begin{eqnarray}
\left.\frac{r_{\rm Ia}}{r_{\rm s}}\right|_{\rm v_{w}^{\rm Ia}} &\approx& 12.4\eta_{0.02} M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4)^{-1}(\tau_{\star}/t_{\rm h})^{1.68} \nonumber \\
&\approx& 12.4M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4)^{-1}(\tau_{\star}/t_{\rm h})^{0.38}
\label{eq:rIars}
\end{eqnarray}
If $r_{\rm Ia} \gg r_{\rm s}$, then Ia heating can only be approximated as a steady heating source near the stagnation radius for extremely massive SMBHs with $M_{\bullet} \gtrsim 10^9 \Msun$ or in the case of a very young stellar population ($\tau_{\star} \ll t_{\rm h}$).  On the other hand, equation (\ref{eq:rIars}) may underestimate the importance of Ia heating because $r_{\rm Ia}/r_{\rm s}$ will be substantially smaller than its value estimated using $v_{w}^{\rm Ia}$ if the heating rate excluding SN Ia is less than this.

>>>>

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other
properties of the galactic nucleus informs key questions related to
the co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{Heckman&Best14}).  SMBH growth
in the low redshift universe is dominated by low mass black holes,
$M_{\bullet} \lesssim 10^{8}M_{\odot}$ (e.g.~\citealt{Heckman+04}), a
fact which is often attributed to the general trend of `cosmic
down-sizing' resulting from hierarchical structure growth (e.g,
\citealt{Gallo+08}).  However, the physical processes involved in the
growth of low mass blacks may be distinct from those operating at
higher SMBH masses and accretion rates in quasars.  A key question
here is whether SMBHs grow primarily by accreting gas fed in directly
from large scales, or whether significant growth results also from
internal, secular processes such as local stellar mass loss.

@@

One observes a similarity between equation (\ref{eq:mdot_gas}) and the Bondi formula (eq.~[\ref{eq:bondi}]) if $r_{\rm B}$ is replaced by $\rs$.  Indeed, when the stagnation radius resides interior to the sphere of influence we find that (eq.~[\ref{eq:stag_simple}])
\begin{align}
  \rs\approx\frac{7}{2}\frac{G \Mbh}{n \vwO^2} \approx \frac{7}{6n}r_{\rm B},
  \label{eq:rs_simple}
\end{align}
where the second equality makes use of equation (\ref{eq:Tanalytic}).

@@

To within a constant of order unity, equation (\ref{eq:cooling2}) also equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3n kT/2)/\dot{q}_{\rm rad}$ to the free-fall time $t_{\rm ff}$ at the stagnation radius, as can be derived using the equality 
\begin{align}
\rho|_{r_{\rm s}}=\frac{3 q\tff|_{r_{\rm s}}}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations
\eqref{eq:mdot_analytic},\eqref{eq:mdot_gas}, and
\eqref{eq:rhostarrs}.  \citet{McCourt+12} argue cooling instability
develops in a hydrostatic atmosphere if $t_{\rm cool} \ll t_{\rm ff}$,
so equation (\ref{eq:cooling2}) represents a good proxy for
instability in this case as well.

@@

In Appendix \ref{app:windheat} we calculate the wind heating rate from
stellar winds, $v_{\rm w}^{\star}$, and the mass loss parameter,
$\eta$ (eq.~[\ref{eq:q}]), as a function of age, $\tau_{\star}$, of a
stellar population which is formed impulsively (Fig.~\ref{fig:vwimp}).
At the earliest times ($\tau_{\star} \lesssim 10^{7}$ yr), the wind
heating rate exceeds 1000 km s$^{-1}$, while much later ($\tau_{\star}
\sim t_{\rm h}$) stellar wind heating dominated by main sequence winds
is much lower, $v^{\star}_{\rm w} \sim 50-100 $ km s$^{-1}$.  We show
in $\S\ref{sec:combined}$ that for the case of quasi-continueous star
formation representative of the average star formation history of loss
mass galaxies, the stellar wind heating rate produced by stellar winds
can also be significant, $v_{\rm w}^{\star} \gtrsim 1000$ km s$^{-1}$.
Stellar winds thus contribute a potentially important source of both
energy and mass for the CNM.

@@

Assuming that $v_{w}^{\rm Ia} \gg \sigma$ then by substituting
$v_{w}^{\rm Ia}$ (eq.~\ref{eq:vIa}) into equation (\ref{eq:rs_simple})
for the stagnation radius, we find that

\begin{eqnarray}
  \left.\frac{r_{\rm Ia}}{r_{\rm s}}\right|_{\rm v_{w}^{\rm Ia}}
  &\approx& 12\eta_{0.02}^{-1} M_{\bullet,8}^{-1.1}(\epsilon_{\rm
    Ia}/0.4) (\tau_{\star}/t_{\rm h})^{-0.56} \nonumber \\
  &\approx& 12 M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4) (\tau_{\star}/t_{\rm h})^{0.74}
\label{eq:rIars}
\end{eqnarray}
{\bf AG-Note the different scaling relations here.}
If $r_{\rm Ia} \gg r_{\rm s}$, then Ia heating can only be
approximated as a steady heating source near the stagnation radius for
extremely massive SMBHs with $M_{\bullet} \gtrsim 10^9 \Msun$ or in
the case of a very young stellar population ($\tau_{\star} \ll t_{\rm
h}$).  On the other hand, equation (\ref{eq:rIars}) may underestimate
the importance of Ia heating because $r_{\rm Ia}/r_{\rm s}$ will be
substantially smaller than its value estimated using $v_{w}^{\rm Ia}$
if the heating rate excluding SN Ia is less than this.
